---
title: "MachineLearningProject"
author: "Nur Sharmini"
date: "November 21, 2015"
output: html_document
---

## Introduction
Jawbone Up, Nike FuelBand and Fitbit are devices which is possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise. 

## Data Processing

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)

```

### 1. Read Data

```{r}
train <- read.csv("F:/R_WD/M8/data/pml-training.csv")
test <- read.csv("F:/R_WD/M8/data/pml-testing.csv")
dim(train)

```

```{r}
dim(test)
```

Training dataset has 19622 observations and 160 variables, while the testing data set contains 20 observations and 160 variables. "classe" variable in the training set is the outcome to predict. 

### 2. Clean Data
Do cleaning data for missing values and remove features that are not in the submit set.

```{r}
# Clean data containing NA.
train <- train[, colSums(is.na(train)) == 0]
test <- test[, colSums(is.na(test)) == 0]
# clean data from features not from the sensors
trainC <- train[, grepl("X|user_name|timestamp|window|^max|^min|^ampl|^var|^avg|^stdd|^ske|^kurt", colnames(train)) == F]
testC <- test[, grepl("X|user_name|timestamp|window|^max|^min|^ampl|^var|^avg|^stdd|^ske|^kurt", 
    colnames(test)) == F]
dim(trainC)
```

```{r}
dim(testC)
```

Now the cleaned data contains 53 variables down from 160 varaibles earlier

#### 3. Slice the data
Split the cleaned training set into a pure training data set (70%) and a validation data set (30%). We will use the validation data set to conduct cross validation in future steps.

```{r}
set.seed(25000) # For reproducibile purpose
inTrain <- createDataPartition(trainC$classe, p=0.70, list=F)
trainX <- trainC[inTrain, ]
validX <- trainC[-inTrain, ]
```

## Data Modeling

Use a predictive model for activity recognition using Random Forest algorithm. We can use 10-fold cross validation when applying the algorithm

```{r}
library(e1071)
controlRf <- trainControl(method="cv", 10)
modelRf <- train(classe ~ ., data=trainX, method="rf", trControl=controlRf, ntree=100, allowParallel=TRUE)
modelRf

```

Predict this modeling on validation data set

```{r}
predictRf <- predict(modelRf, validX)
confusionMatrix(validX$classe, predictRf)
```

Check Accuracy

```{r}
accuracy <- postResample(predictRf, validX$classe)
accuracy

```

Accuracy of this prediction model is 99.18%.

Out of Sample Error is calculated here

```{r}
error <- 1 - as.numeric(confusionMatrix(validX$classe, predictRf)$overall[1])
error
```

Out of Sample Error is 0.82%

## Predicting for Test Data Set

Apply the model to the original testing data set downloaded from the data source. Generate "problem_id_x.txt" Files for the assignments. These generated individual files are stored in working directory.

```{r}
Testpredict <- predict(modelRf, testC, type = "raw")
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(Testpredict)

```

Loaded these files as per Problem_id number into Project Assignment and obtained the correct results for all 20 cases.

## Appendix : Tree Visualization

```{r}
treeModel <- rpart(classe ~ ., data=trainX, method="class")
prp(treeModel) # plot of Tree Model

```

